---
title: "Take-home_Ex02 - Fable"
date: "March 14, 2026"
date-modified: "`r Sys.Date()`"
---

# Take-home Exercise 2: fable

## Data Preparation

### Import Packages
Similar to **modeltime** pipeline, we need to load the necessary packages using `pacman::p_load()`. The key difference is tidymodels packages are replaced with tidyverts packages (`tsibble`, `fable`, `fable.prophet`, `fabletools`, and `feasts`).
```{r}
pacman::p_load(readxl, tidyverse, tsibble, fable, fable.prophet, fabletools, feasts, lubridate)
```

### Import Data
```{r}
cpi_data_raw <- read_excel("data/SG_CPI_012020_012026_Use.xlsx")
```


## Data Wrangling
### Define Hierarchy Map
```{r}
hier_map <- tribble(
  ~Category,                                                                    ~Level, ~Parent,

  # Level 1 - Root
  "Consumer Price Index (CPI)",                                                  1,      NA,

  # Level 2 - Major Categories
  "CPI: Food",                                                                   2,      "Consumer Price Index (CPI)",
  "CPI: Clothing and Footwear (C&F)",                                           2,      "Consumer Price Index (CPI)",
  "CPI: Housing & Utilities",                                                   2,      "Consumer Price Index (CPI)",
  "CPI: Household Durables & Services (HDS)",                                   2,      "Consumer Price Index (CPI)",
  "CPI: Health",                                                                2,      "Consumer Price Index (CPI)",
  "CPI: Transport",                                                             2,      "Consumer Price Index (CPI)",
  "CPI: Information & Communication (InfoComm)",                                2,      "Consumer Price Index (CPI)",
  "CPI: Recreation, Sport & Culture (RSC)",                                     2,      "Consumer Price Index (CPI)",
  "CPI: Education",                                                             2,      "Consumer Price Index (CPI)",
  "CPI: Miscellaneous Goods & Services (MG&S)",                                 2,      "Consumer Price Index (CPI)",

  # Level 3 - Food sub-categories
  "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",                    3,      "CPI: Food",
  "CPI: Food & Beverage Serving Services (FBSS)",                               3,      "CPI: Food",

  # Level 3 - C&F children
  "CPI: C&F: Clothing",                                                         3,      "CPI: Clothing and Footwear (C&F)",
  "CPI: C&F: Footwear",                                                         3,      "CPI: Clothing and Footwear (C&F)",

  # Level 3 - Housing & Utilities children
  "CPI: Housing & Utilities: Accommodation",                                    3,      "CPI: Housing & Utilities",
  "CPI: Housing & Utilities: Utilities & Other Fuels",                          3,      "CPI: Housing & Utilities",

  # Level 3 - HDS children
  "CPI: HDS: Furniture & Furnishings",                                          3,      "CPI: Household Durables & Services (HDS)",
  "CPI: HDS: Household Textiles",                                               3,      "CPI: Household Durables & Services (HDS)",
  "CPI: HDS: Household Appliances",                                             3,      "CPI: Household Durables & Services (HDS)",
  "CPI: HDS: Glassware, Tableware & Household Utensils",                        3,      "CPI: Household Durables & Services (HDS)",
  "CPI: HDS: Tools & Equipment For House & Garden",                             3,      "CPI: Household Durables & Services (HDS)",
  "CPI: HDS: Household Services & Supplies (HSS)",                              3,      "CPI: Household Durables & Services (HDS)",

  # Level 3 - Health children
  "CPI: Health: Medicines & Health Products",                                   3,      "CPI: Health",
  "CPI: Health: Outpatient Services",                                           3,      "CPI: Health",
  "CPI: Health: Inpatient Services",                                            3,      "CPI: Health",
  "CPI: Health: Other Health Services (OHS)",                                   3,      "CPI: Health",
  "CPI: Health: Health Insurance",                                              3,      "CPI: Health",

  # Level 3 - Transport children
  "CPI: Transport: Private Transport",                                          3,      "CPI: Transport",
  "CPI: Transport: Land Transport Services",                                    3,      "CPI: Transport",
  "CPI: Transport: Other Transport Services",                                   3,      "CPI: Transport",
  "CPI: Transport: Transport Services Of Goods",                                3,      "CPI: Transport",

  # Level 3 - InfoComm children
  "CPI: InfoComm: Information & Communication Equipment",                       3,      "CPI: Information & Communication (InfoComm)",
  "CPI: InfoComm: Software Excl Games",                                         3,      "CPI: Information & Communication (InfoComm)",
  "CPI: InfoComm: InfoComm Services",                                           3,      "CPI: Information & Communication (InfoComm)",

  # Level 3 - RSC children
  "CPI: RSC: Recreational & Cultural Goods",                                    3,      "CPI: Recreation, Sport & Culture (RSC)",
  "CPI: RSC: Other Recreational Goods (ORG)",                                   3,      "CPI: Recreation, Sport & Culture (RSC)",
  "CPI: RSC: Garden Products & Pets",                                           3,      "CPI: Recreation, Sport & Culture (RSC)",
  "CPI: RSC: Recreational Services",                                            3,      "CPI: Recreation, Sport & Culture (RSC)",
  "CPI: RSC: Cultural Services",                                                3,      "CPI: Recreation, Sport & Culture (RSC)",
  "CPI: RSC: Newspapers, Books & Stationery (NBS)",                             3,      "CPI: Recreation, Sport & Culture (RSC)",
  "CPI: RSC: Holiday Expenses",                                                 3,      "CPI: Recreation, Sport & Culture (RSC)",

  # Level 3 - Education children
  "CPI: Education: General, Vocational & Higher Education (GVH Edu)",           3,      "CPI: Education",
  "CPI: Education: Private Tuition & Other Educational Courses",                3,      "CPI: Education",
  "CPI: Education: School Textbooks & Study Guides",                            3,      "CPI: Education",

  # Level 3 - MG&S children
  "CPI: MG&S: Personal Care",                                                   3,      "CPI: Miscellaneous Goods & Services (MG&S)",
  "CPI: MG&S: Alcoholic Drinks & Tobacco",                                      3,      "CPI: Miscellaneous Goods & Services (MG&S)",
  "CPI: MG&S: Personal Effects",                                                3,      "CPI: Miscellaneous Goods & Services (MG&S)",
  "CPI: MG&S: Social Services",                                                 3,      "CPI: Miscellaneous Goods & Services (MG&S)",
  "CPI: MG&S: Other Miscellaneous Services",                                    3,      "CPI: Miscellaneous Goods & Services (MG&S)",

  # Level 4 - FFBSS children (deepest level)
  "CPI: FFBSS: Rice & Cereals",                                                 4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",
  "CPI: FFBSS: Meat",                                                           4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",
  "CPI: FFBSS: Fish & Seafood",                                                 4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",
  "CPI: FFBSS: Milk, Other Dairy & Eggs",                                       4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",
  "CPI: FFBSS: Oils & Fats",                                                    4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",
  "CPI: FFBSS: Fruits & Nuts (FN)",                                             4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",
  "CPI: FFBSS: Vegetables",                                                     4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",
  "CPI: FFBSS: Sugar, Confectionery & Desserts (SCD)",                          4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",
  "CPI: FFBSS: Ready-Made & Oth. Food Nec",                                     4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",
  "CPI: FFBSS: Non-Alcoholic Beverages",                                        4,      "CPI: Food Excl Food & Beverage Serving Services (FFBSS)",

  # Level 4 - FBSS children (deepest level)
  "CPI: FBSS: Restaurants, Cafes & Pubs",                                       4,      "CPI: Food & Beverage Serving Services (FBSS)",
  "CPI: FBSS: Fast Food",                                                       4,      "CPI: Food & Beverage Serving Services (FBSS)",
  "CPI: FBSS: Hawker Food Incl. Food Courts (HF&FC)",                           4,      "CPI: Food & Beverage Serving Services (FBSS)",
  "CPI: FBSS: Oth Catering Services, Incl Vending Machines",                    4,      "CPI: Food & Beverage Serving Services (FBSS)"
)
```

### Reshape Data to Long
```{r}
cpi_data_long <- cpi_data_raw %>%
  pivot_longer(
    cols = -Date,
    names_to = "Category",
    values_to = "CPI_Value"
  ) %>%
  mutate(Date = floor_date(as.Date(Date), "month")) %>%
  left_join(hier_map, by = "Category")
```

### Identify Leaf Nodes
```{r}
leaf_categories <- hier_map %>%
  filter(!Category %in% na.omit(hier_map$Parent)) %>%
  pull(Category)
```

### Filter to Leaf Nodes
While **modeltime** uses `pad_by_time()` to fill time gaps, **fable uses** `fill_gaps()` on a tsibble object. 

Another difference is that an additional step of defining `mutate(Date = yearmonth(Date))` is needed to convert date column into **tsibble** native monthly index type to explicitly tell **tsibble** that the data is monthly. 
```{r}
cpi_tsibble <- cpi_data_raw %>%
  pivot_longer(
    cols      = -Date,
    names_to  = "Category",
    values_to = "CPI_Value"
  ) %>%
  mutate(Date = floor_date(as.Date(Date), "month")) %>%
  left_join(hier_map, by = "Category") %>%
  filter(Category %in% leaf_categories) %>%
  select(Date, Category, CPI_Value) %>%
  mutate(Date = yearmonth(Date)) %>%   # convert AFTER floor_date, so tsibble knows it's monthly
  as_tsibble(key = Category, index = Date) %>%
  fill_gaps() %>%
  group_by_key() %>%
  mutate(CPI_Value = zoo::na.approx(CPI_Value, na.rm = FALSE)) %>%
  ungroup()
```

### Missingness Check
```{r}
cpi_tsibble %>% filter(is.na(CPI_Value))
```

## Time Series Forecasting
### Train-test Split
Similar to **modeltime**, the last 12 months are reserved as a hold-out period (test set). The key difference is that instead of using `split_nested_timeseries()`, `filter()` is used. Since `Date` is a `yearmonth` object, we can do direct integer subtraction instead of using lubridate `%m-%`, which is incompatible with `yearmonth`.

```{r}
cutoff_date <- max(cpi_tsibble$Date) - 12

train_tbl <- cpi_tsibble %>% filter(Date <= cutoff_date)
test_tbl <- cpi_tsibble %>% filter(Date > cutoff_date)
```

### Fit models on Train Set
While the **modeltime** pipeline requires us to explicitly define workflow pairings for each model with preprocessing recipe, **fable's** `model()` function is more straight forward. 

All three models (ETS, ARIMA, Prophet) are fitted across all 51 leaf categories in a single cell. This is possible because tsibble's `key` variable is capable to handle multi-series structure, eliminating the need for `nest_timeseries()` and `modeltime_nested_fit()`. 

However, **ARIMA-XGBoost is not available in fable**.

```{r}
fit_train <- train_tbl %>%
  model(
    ets = ETS(CPI_Value),
    arima = ARIMA(CPI_Value),
    prophet = prophet(CPI_Value)
  )
```


### Select the Best Model
Simliar to `modeltime_nested_select_best()`, the best model per category is selected by minimising RMSE on the test set. 

However, **fable** does not have built-in function that is similar to `modeltime_nested_select_best()`. Hence, this is done manually by computing accuracy metrics with `accuracy()`, using `slice_min(RMSE)` and `semi_join()` to identify and retain the best model for each category.

```{r}
acc_tbl <- fit_train %>%
  forecast(h = 12) %>%
  accuracy(cpi_tsibble)

best_model <- acc_tbl %>%
  group_by(Category) %>%
  slice_min(RMSE, n = 1) %>%
  ungroup() %>%
  select(Category, .model)
```

### Refit on Full Dataset
```{r}
fit_full <- cpi_tsibble %>%
  model(
    ets = ETS(CPI_Value),
    arima = ARIMA(CPI_Value),
    prophet = prophet(CPI_Value)
  )
```

### Generate 12-month Forecast
```{r}
leaf_fc <- fit_full %>%
  forecast(h = 12) %>%
  semi_join(best_model, by = c("Category", ".model")) %>%
  as_tibble() %>%
  mutate(Date = as.Date(Date)) %>%
  select(Date, Category, .value = .mean)
```

### Define Weights
```{r}
weights_tbl <- tribble(
  ~Category,                                                                    ~Weight,

  # Level 4 - FFBSS leaf nodes
  "CPI: FFBSS: Rice & Cereals",                                                  122,
  "CPI: FFBSS: Meat",                                                            101,
  "CPI: FFBSS: Fish & Seafood",                                                   79,
  "CPI: FFBSS: Milk, Other Dairy & Eggs",                                         64,
  "CPI: FFBSS: Oils & Fats",                                                      12,
  "CPI: FFBSS: Fruits & Nuts (FN)",                                               79,
  "CPI: FFBSS: Vegetables",                                                       82,
  "CPI: FFBSS: Sugar, Confectionery & Desserts (SCD)",                            27,
  "CPI: FFBSS: Ready-Made & Oth. Food Nec",                                       44,
  "CPI: FFBSS: Non-Alcoholic Beverages",                                          41,

  # Level 4 - FBSS leaf nodes
  "CPI: FBSS: Restaurants, Cafes & Pubs",                                        585,
  "CPI: FBSS: Fast Food",                                                         85,
  "CPI: FBSS: Hawker Food Incl. Food Courts (HF&FC)",                            707,
  "CPI: FBSS: Oth Catering Services, Incl Vending Machines",                      14,

  # Level 3 - leaf nodes for all other branches
  "CPI: C&F: Clothing",                                                          129,
  "CPI: C&F: Footwear",                                                           36,
  "CPI: Housing & Utilities: Accommodation",                                    2656,
  "CPI: Housing & Utilities: Utilities & Other Fuels",                           282,
  "CPI: HDS: Furniture & Furnishings",                                           128,
  "CPI: HDS: Household Textiles",                                                 17,
  "CPI: HDS: Household Appliances",                                              101,
  "CPI: HDS: Glassware, Tableware & Household Utensils",                          17,
  "CPI: HDS: Tools & Equipment For House & Garden",                                5,
  "CPI: HDS: Household Services & Supplies (HSS)",                               279,
  "CPI: Health: Medicines & Health Products",                                     94,
  "CPI: Health: Outpatient Services",                                            428,
  "CPI: Health: Inpatient Services",                                             238,
  "CPI: Health: Other Health Services (OHS)",                                     23,
  "CPI: Health: Health Insurance",                                               225,
  "CPI: Transport: Private Transport",                                           906,
  "CPI: Transport: Land Transport Services",                                     262,
  "CPI: Transport: Other Transport Services",                                    131,
  "CPI: Transport: Transport Services Of Goods",                                   8,
  "CPI: InfoComm: Information & Communication Equipment",                         91,
  "CPI: InfoComm: Software Excl Games",                                            3,
  "CPI: InfoComm: InfoComm Services",                                            287,
  "CPI: RSC: Recreational & Cultural Goods",                                       9,
  "CPI: RSC: Other Recreational Goods (ORG)",                                     33,
  "CPI: RSC: Garden Products & Pets",                                             27,
  "CPI: RSC: Recreational Services",                                             144,
  "CPI: RSC: Cultural Services",                                                  52,
  "CPI: RSC: Newspapers, Books & Stationery (NBS)",                               20,
  "CPI: RSC: Holiday Expenses",                                                  310,
  "CPI: Education: General, Vocational & Higher Education (GVH Edu)",            391,
  "CPI: Education: Private Tuition & Other Educational Courses",                 184,
  "CPI: Education: School Textbooks & Study Guides",                               4,
  "CPI: MG&S: Personal Care",                                                    211,
  "CPI: MG&S: Alcoholic Drinks & Tobacco",                                        58,
  "CPI: MG&S: Personal Effects",                                                  62,
  "CPI: MG&S: Social Services",                                                   32,
  "CPI: MG&S: Other Miscellaneous Services",                                      75
)
```

### Aggregate Forecast
```{r}
aggregate_bottom_up <- function(forecasts, hier_map, weights_tbl) {
  all_levels <- forecasts %>%
    left_join(weights_tbl, by = "Category")

  levels_to_agg <- hier_map %>%
    filter(!is.na(Parent)) %>%
    arrange(desc(Level)) %>%
    pull(Level) %>%
    unique()

  for (lvl in levels_to_agg) {
    child_nodes <- hier_map %>%
      filter(Level == lvl) %>%
      select(Category, Parent)

    parent_fc <- all_levels %>%
      inner_join(child_nodes, by = "Category") %>%
      group_by(Date, Category = Parent) %>%
      summarise(
        .value = sum(.value * Weight) / sum(Weight),
        Weight = sum(Weight),
        .groups = "drop"
      )

    all_levels <- bind_rows(all_levels, parent_fc) %>%
      distinct(Date, Category, .keep_all = TRUE)
  }
  return(all_levels)
}

reconciled_forecasts <- aggregate_bottom_up(leaf_fc, hier_map, weights_tbl)
```

### Visualisation of Actual and Forecasted CPI

#### Plot by CPI Level
```{r}
actuals <- cpi_data_long %>%
  select(Date, Category, CPI_Value) %>%
  mutate(.key = "Actual")

forecasts_plot <- reconciled_forecasts %>%
  select(Date, Category, CPI_Value = .value) %>%
  mutate(.key = "Forecast")

# Plot by hierarchy level
plot_by_level <- function(lvl) {
  cats <- hier_map %>% filter(Level == lvl) %>% pull(Category)

  bind_rows(actuals, forecasts_plot) %>%
    filter(Category %in% cats) %>%
    ggplot(aes(x = Date, y = CPI_Value, color = .key, linetype = .key)) +
    geom_line(linewidth = 0.7) +
    scale_color_manual(values = c("Actual" = "#378ADD", "Forecast" = "#EF9F27")) +
    scale_linetype_manual(values = c("Actual" = "solid", "Forecast" = "dashed")) +
    facet_wrap(~ Category, scales = "free_y", ncol = 3) +
    labs(title = paste("Singapore CPI — Level", lvl, "| Actual vs Forecast"),
         x = NULL, y = "CPI Value", color = NULL, linetype = NULL) +
    theme_minimal(base_size = 11) +
    theme(
      strip.text       = element_text(size = 9, face = "bold"),
      axis.text.x      = element_text(angle = 45, hjust = 1),
      panel.grid.minor = element_blank(),
      legend.position  = "bottom"
    )
}

plot_by_level(1)
plot_by_level(2)
plot_by_level(3)
plot_by_level(4)
```

#### Plot by Parent Node
```{r}
plot_by_parent <- function(parent_cat) {
  cats <- hier_map %>% filter(Parent == parent_cat) %>% pull(Category)
  n    <- length(cats)

  bind_rows(actuals, forecasts_plot) %>%
    filter(Category %in% cats) %>%
    ggplot(aes(x = Date, y = CPI_Value, color = .key, linetype = .key)) +
    geom_line(linewidth = 0.7) +
    scale_color_manual(values = c("Actual" = "#378ADD", "Forecast" = "#EF9F27")) +
    scale_linetype_manual(values = c("Actual" = "solid", "Forecast" = "dashed")) +
    facet_wrap(~ Category, scales = "free_y", ncol = min(3, n)) +
    labs(
      title    = paste("Sub-categories of:", parent_cat),
      subtitle = "Actual vs Forecast",
      x = NULL, y = "CPI Value", color = NULL, linetype = NULL
    ) +
    theme_minimal(base_size = 11) +
    theme(
      strip.text       = element_text(size = 9, face = "bold"),
      axis.text.x      = element_text(angle = 45, hjust = 1),
      panel.grid.minor = element_blank(),
      legend.position  = "bottom"
    )
}

level2_parents <- hier_map %>% filter(Level == 2) %>% pull(Category)
walk(level2_parents, ~ print(plot_by_parent(.x)))
```
## Saving Model Metrics for Comparison
```{r}
# Save fable accuracy and time
accuracy_fable <- fit_train %>%
  forecast(h = 12) %>%
  accuracy(cpi_tsibble) %>%
  semi_join(best_model, by = c("Category", ".model")) %>%
  select(Category, .model, RMSE, MAE, MAPE) %>%
  mutate(framework = "fable")

time_fable <- system.time({
  train_tbl %>%
    model(
      ets     = ETS(CPI_Value),
      arima   = ARIMA(CPI_Value),
      prophet = prophet(CPI_Value)
    )
})

saveRDS(accuracy_fable, "data/accuracy_fable.rds")
saveRDS(time_fable,     "data/time_fable.rds")
```

